---
title: "0-4 vs 5-14 TB Difference"
output: html_notebook
---

Aim: Create two data frames containing information about the TB burden estimate, reported cases, absolute difference and percentage difference for children aged 0-4 and 5-14 years.

Loading the data 

```{r}
data <- read_csv("TBburden.csv")
head(data)
colnames(data)
data1 <- data[ , c("country", "year", "age_group", "sex", "best", "lo", "hi")]
head(data1)
data2 <- unite(data1, "range", "lo", "hi", sep = "-")
```
Creating a data frame for the 0-4 only.

```{r}
data3 <- data2[data2$age_group == "0-4", ]
head(data3)
```
Checking that each country has a value for f and m.
```{r}
data3$sex <- as.factor(data3$sex)
str(data3$sex)
identical(sum(data3$sex == "f"), sum(data3$sex == "m"))
str(data3$country)
data3$country <- as.factor(data3$country)
str(data3$country)
```

Combining the male and female data into a new column containing total burnde for that age group.

```{r}
summed_data <- aggregate(x = data3$best, 
                         by = list(data3$country), 
                         FUN = sum)
colnames(summed_data)[colnames(summed_data) == "x"] <- "0-4"
colnames(summed_data)[colnames(summed_data) == "Group.1"] <- "Country" 
head(summed_data)
burden_04 <- summed_data
```

Creating data frame for 5-14 only.
```{r}
data_514 <- data2[data2$age_group == "5-14", ]
head(data_514)
```
Checking that each country has a value for m and f 

```{r}
data_514$sex <- as.factor(data_514$sex)
str(data_514$sex)
identical(sum(data_514$sex == "f"), sum(data_514$sex == "m"))
```

Creating new table with total case burden for 5-14 years for each country.

```{r}
summed_data514 <- aggregate(x = data_514$best, 
                         by = list(data_514$country), 
                         FUN = sum)
head(summed_data514)

colnames(summed_data514)[colnames(summed_data514) == "Group.1"] <- "Country"
colnames(summed_data514)[colnames(summed_data514) == "x"] <- "5-14"

burden_514 <- summed_data514

```

Checking that countries match.

```{r}
typeof(burden_514$Country)
typeof(burden_04$Country)

burden_04$Country <- as.character(burden_04$Country)

identical(burden_514$Country, burden_04$Country)
```

Creating final data frame with both sets of data.

```{r}
burden_agesplit <- cbind(burden_04, burden_514$`5-14`)
colnames(burden_agesplit)[colnames(burden_agesplit) == "burden_514$`5-14`"] <- "5-14"
head(burden_agesplit)
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

