---
title: "Helminth Infections and TB"
output: html_notebook
---

Importing PDF data about helminth infection rates. 

```{r}
helminth <- read_excel("helminth.xlsx", skip = 3)
helminth1 <- helminth[ , c(1:4, 6:7)]
colnames(helminth1) <- c("Country", "Total_IUs", "STH_Prevalence", "Mod_Sev_Prev", "High_Prev_Proportion", "IUs_above_target")
```

Filtering the case difference data 

```{r}
setwd("~/Desktop/AFP/Original Data")
TBburden <- read_csv("TBburden.csv")

#isolating the important data
firstfilter <- TBburden %>% select(country, iso3, age_group, sex, best, lo, hi)

#focusing on age group 0-14
secondfilter <- filter(firstfilter, age_group == "0-14") 

#combining the range into one column
burdensummary <- unite(secondfilter, "range", lo, hi, sep = "-") 

#combining sexes
allcases <- burdensummary %>% filter(sex == "a") 

#Cutting out columns 
final_burden <- allcases[, c("country", "iso3", "best")]

#Reported data 
ISO_final <- ISO_combined[, c("country", "iso3", "Cases")]

#Length of the data sets 
length(ISO_final$country)
length(final_burden$country)

#Matching the data 
matched <- match(ISO_final$country, final_burden$country)
matched_burden <- final_burden[c(matched), ]
identical(ISO_final$country, matched_burden$country)

#Combining the data frames 
combined_burden <- cbind(matched_burden, ISO_final$Cases)
colnames(combined_burden)[colnames(combined_burden) == "country"] <- "Country"
colnames(combined_burden)[colnames(combined_burden) == "iso3"] <- "ISO3"
colnames(combined_burden)[colnames(combined_burden) == "best"] <- "Estimated"
colnames(combined_burden)[colnames(combined_burden) == "ISO_final$Cases"] <- "Reported"

#Adding the difference columns
burden_diff <- mutate(combined_burden, "Difference" = combined_burden$Estimated - combined_burden$Reported)
burden_diff_perc <- mutate(combined_burden, "Perc_Diff" = (combined_burden$Reported/combined_burden$Estimated)*100)

#Remove NaN 
burden_nona <- na.omit(burden_diff_perc)
```

Matching up the case difference data with the helminth data

```{r}
burden_data <- burden_nona 
helminth_data <- helminth1
match_hel <- match(helminth1$Country, burden_data$ISO3)
match_nona <- na.omit(match_hel)
filter_burden <- burden_data[c(match_nona), ]
match_hel2 <- match(filter_burden$ISO3, helminth_data$Country)
filter_hel <- helminth_data[c(match_hel2), ]
identical(filter_burden$ISO3, filter_hel$Country)
```
Combining helminth data with case data

```{r}
bound_hel <- cbind(filter_burden, filter_hel[, c(2:6)])
```

Adding a column for absolute difference 

```{r}
bound_hel2 <- mutate(bound_hel, "Absolute" = Estimated - Reported)
```

Creating graph of percentage reported TB cases gainst STH prevalence

```{r}
graph1 <- ggplot(bound_hel2, aes(x = STH_Prevalence, y = Perc_Diff)) + geom_point() + geom_smooth()
```

Creating graph of percentage reported TB cases against percentage of IUs that are above the target range 

```{r}
target <- mutate(bound_hel2, "Target_Perc" = (IUs_above_target/Total_IUs)*100)
graph2 <- ggplot(target, aes(x = Target_Perc, y = Perc_Diff)) + geom_point() + geom_smooth()
```


NEXT STEPS - add the burden data as a column and add difference and proportions as a column. Then make a few graphs. 


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

