---
title: "Vaccine-Preventable Diseases"
output: html_notebook
---

AIM: Create a data frame and a visualisation looking at whether there is a correlation between 
rates of vaccine-preventable diseases and difference in TB burden/notifications 

Extracting the data from multiple csv files 
```{r}
extract.data <- function(filename){
        x <- read_csv(filename, skip = 1)
        return(x) }

diphtheria <- extract.data("Diphtheria.csv")
measles <- extract.data("Measles.csv")
mumps <- extract.data("Mumps.csv")
pertussis <- extract.data("Pertussis.csv")
polio <- extract.data("Polio.csv")
rubella <- extract.data("Rubella.csv")
tetanus <- extract.data("Tetanus.csv")

```

Isolating data for 2019 only

```{r}
data_2019 <- function(disease) {
        y <- disease %>% select("Country", "2019")
        return(y) }

diphtheria_2019 <- data_2019(diphtheria)
measles_2019 <- data_2019(measles)
mumps_2019 <- data_2019(mumps)
pertussis_2019 <- data_2019(pertussis)
polio_2019 <- data_2019(polio)
rubella_2019 <- data_2019(rubella)
tetanus_2019 <- data_2019(tetanus)

diseases <- c(diphtheria_2019, measles_2019, mumps_2019, pertussis_2019, 
              polio_2019, rubella_2019, tetanus_2019)
```

Testing columns dimensions and to see which columns are identical 

```{r}
dim(diphtheria_2019)
dim(measles_2019)
dim(mumps_2019)
dim(pertussis_2019)
dim(polio_2019)
dim(rubella_2019)
dim(tetanus_2019)

all.equal(diphtheria_2019$Country, measles_2019$Country, pertussis_2019$Country, rubella_2019$Country, tetanus_2019$Country)
```

Mumps has the fewest data so the data for countries that match the mumps set will need to be extracted from the other sets. 

```{r}
match.mumps <- function(disease, mumps_2019) {
        z <- disease[match(mumps$Country, disease$Country), ]
        return(z)}

diph_matched <- match.mumps(diphtheria_2019, mumps_2019)
meas_matched <- match.mumps(measles_2019, mumps_2019)
pert_matched <- match.mumps(pertussis_2019, mumps_2019)
polio_matched <- match.mumps(polio_2019, mumps_2019)
rubella_matched <- match.mumps(rubella_2019, mumps_2019)
tet_matched <- match.mumps(tetanus_2019, mumps_2019)

all.equal(diph_matched$Country, meas_matched$Country, pert_matched$Country, polio_matched$Country, rubella_matched$Country, tet_matched$Country, mumps_2019$Country)

```

Combining into a single data frame.

```{r}
bound_vax <- cbind(mumps_2019, diph_matched$`2019`, meas_matched$`2019`, pert_matched$`2019`, 
                   polio_matched$`2019`, rubella_matched$`2019`, tet_matched$`2019`)

colnames(bound_vax) <- c("Country", "Mumps", "Diphtheria", "Measles", 
                                              "Pertussis", "Polio", "Rubella", "Tetanus")

complete_vax <- bound_vax[complete.cases(bound_vax), ]
```

Importing the TB burden/notification difference data. 

```{r}
tb.data <- percent_rep
match_tb <- match(complete_vax$Country, tb.data$Country)
tb.filter <- tb.data[match_tb, ]
tb.filter.nona <- na.omit(tb.filter)
vax.filter.nona <- complete_vax[match(tb.filter.nona$Country, complete_vax$Country), ]
identical(tb.filter.nona$Country, vax.filter.nona$Country)

```

Putting data frames together 

```{r}
vax_tb <- cbind(tb.filter.nona, vax.filter.nona[ , 2:8])
total_vax <- mutate(vax_tb, Total = Mumps + Diphtheria + Measles + Pertussis + Polio + Rubella + Tetanus)
```

First plot 

```{r}
plot(x = total_vax$Total, y = total_vax$`Percent Reported`)
```

Removing burden < 100 cases 

```{r}
filter.vax <- filter(total_vax, Estimated > 100, Total < 30000)
plot(x = filter.vax$Total, y = filter.vax$`Percent Reported`, main = "Vax Preventable Disease vs Percentage Reported TB",xlab = "Vax_Total", ylab = "Percent_Reported_TB")
abline(lm(`Percent Reported` ~ Total, data = filter.vax), col = "red")

```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

